Measurement and Probability in Fuzzy Quantum Theories

### 2017-04-15 prop_iop.tex
prop_iop.tex \section{Expectation Values for IVPMs} Updating...

We may need to understand how to define expectation values of an observable
for quantum IVPMs. Recall an observable is the quantum version of
a random variable. Hence, in order to define expectation values of
an observable, we need to understand the definition of expectation
values of a random variable for an classical IVPM. Since the expectation
value of random variable~$X$ is the integral $\int X\rmd\pmeas$~\cite{Kolmogorov1950},
we need to understand the meaning of integration on IVPMs. In the
rest of the section, we will consider the real-valued and interval-valued
probability measure defined in example~\ref{ex1} and \ref{ex3},
and see how to compute the expectation value of different random variables.
To simplify the discussion later, we will only consider a random variable~$X:\Omega\rightarrow\left[0,1\right]$
which is a step function, i.e., $X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$,
where $E_{i}\in\events$ and $\mathbf{1}_{E}$ is the indicator function
defined by 
\begin{equation} 
\mathbf{1}_{A}(x)=\cases{ 
1 & if $x\in A$ ;\\ 
0 & if $x\notin A$. 
}
\end{equation}

Given a random variable~$X$ and a probability measure~$\pmeas$,
the expectation value of $X$ is 
\begin{eqnarray}
\int X\rmd\pmeas & = & \sum_{i=1}^{N}x_{i}\pmeas\left(E_{i}\right)\textrm{ .}
\end{eqnarray}
Intuitively, the expectation value is weighted average of the possible
outcomes of $X$ with the weight equal to the probability of each
outcome could happen. For example, consider the random variable~$X_{0}$
representing the number of tails. For the probability measure~$\pmeas$
in example~\ref{ex1}, the expectation value of $X_{0}$ is
\begin{eqnarray*}
\int X_{0}\rmd\pmeas & = & 0\cdot\pmeas(\{HH\})+1\cdot\pmeas(\{HT,TH\})+2\cdot\pmeas(\{TT\})\\
 & = & 0\cdot\frac{1}{3}+1\cdot\frac{2}{3}+2\cdot0=\frac{2}{3}\textrm{ .}
\end{eqnarray*}

After we discussed the conventional expectation value, we can start
to discuss the expectation value of IVPM. First, the only source use
the exact same definition as ours is written by Jamison and Lodwick~\cite{JamisonLodwick2004}
so it's reasonable to start from their definition of integration\footnote{Although they said they would explore their integration in more details,
they change their definition of IVPM later. Hence, I don't know if
they have any further development of their integration.}. Their integral of $X$ is defined by 
\begin{eqnarray*}
\int_{E}^{\textrm{JL}}X\rmd\bar{\mu} & = & \left[\inf\set{\sum_{i=1}^{N}x_{i}e_{i}}{e_{i}\in\bar{\mu}\left(E_{i}\right)\textrm{ and }\sum_{i=1}^{N}e_{i}=\min\bar{\mu}\left(E\right)},\right.\\
 &  & \left.\sup\set{\sum_{i=1}^{N}x_{i}e_{i}}{e_{i}\in\bar{\mu}\left(E_{i}\right)\textrm{ and }\sum_{i=1}^{N}e_{i}=\max\bar{\mu}\left(E\right)}\right]\textrm{ .}
\end{eqnarray*}
In this definition, we still work on the weighted average. Since we
don't know the exact probability, we want to average on all possible
allowed probability, which is represented by the requirement on $e_{i}\in\bar{\mu}\left(E_{i}\right)$.
However, I don't have a clear intuition on why they want $\sum_{i=1}^{N}e_{i}=\min\bar{\mu}\left(E\right)$
and $\sum_{i=1}^{N}e_{i}=\max\bar{\mu}\left(E\right)$, and we will
have a more intuitive definition following this line when we move
on to Choquet integral~\cite{Choquet1954,Grabisch2016}. After we
have the definition, we can consider the IVPM~$\bar{\mu}$ in example~\ref{ex3}.
Because $\bar{\mu}\left(\Omega\right)=\left[1,1\right]$, we have
\begin{eqnarray*}
\int_{\Omega}^{\textrm{JL}}X\rmd\bar{\mu} & = & \left[\inf\set{\sum_{i=1}^{N}x_{i}e_{i}}{e_{i}\in\bar{\mu}\left(E_{i}\right)\textrm{ and }\sum_{i=1}^{N}e_{i}=1},\right.\\
 &  & \left.\sup\set{\sum_{i=1}^{N}x_{i}e_{i}}{e_{i}\in\bar{\mu}\left(E_{i}\right)\textrm{ and }\sum_{i=1}^{N}e_{i}=1}\right]\\
 & = & \left[\inf\set{e_{1}+2e_{2}}{e_{0}\in\bar{\mu}\left(\{HH\}\right)\textrm{ , }e_{1}\in\bar{\mu}\left(\{HT,TH\}\right)\textrm{ , }e_{2}\in\bar{\mu}\left(\{TT\}\right)\textrm{ , and }e_{0}+e_{1}+e_{2}=1},\right.\\
 &  & \left.\sup\set{e_{1}+2e_{2}}{e_{0}\in\bar{\mu}\left(\{HH\}\right)\textrm{ , }e_{1}\in\bar{\mu}\left(\{HT,TH\}\right)\textrm{ , }e_{2}\in\bar{\mu}\left(\{TT\}\right)\textrm{ , and }e_{0}+e_{1}+e_{2}=1}\right]\\
 & = & \left[\inf\set{e_{1}+2e_{2}}{0\le e_{0}\le\frac{1}{2}\textrm{ , }\frac{1}{2}\le e_{1}\le1\textrm{ , }e_{2}=0\textrm{ , and }e_{0}+e_{1}+e_{2}=1},\right.\\
 &  & \left.\sup\set{e_{1}+2e_{2}}{0\le e_{0}\le\frac{1}{2}\textrm{ , }\frac{1}{2}\le e_{1}\le1\textrm{ , }e_{2}=0\textrm{ , and }e_{0}+e_{1}+e_{2}=1}\right]\\
 & = & \left[\inf\set{e_{1}}{0\le e_{0}\le\frac{1}{2}\textrm{ , }\frac{1}{2}\le e_{1}\le1\textrm{ , and }e_{0}+e_{1}=1},\right.\\
 &  & \left.\sup\set{e_{1}}{0\le e_{0}\le\frac{1}{2}\textrm{ , }\frac{1}{2}\le e_{1}\le1\textrm{ , and }e_{0}+e_{1}=1}\right]\\
 & = & \left[\frac{1}{2},1\right]\textrm{ .}
\end{eqnarray*}

\yutsung{{Put the definition of }Choquet integral, other meta-discourse
here...}

>>>>

We may need to understand how to define expectation values of an observable
for quantum IVPMs. Recall an observable is the quantum version of
a random variable. Hence, in order to define expectation values of
an observable, we need to understand the definition of expectation
values of a random variable for an classical IVPM. Since the expectation
value of random variable~$X$ is the integral $\int X\rmd\pmeas$~\cite{Kolmogorov1950},
we need to understand the meaning of integration on IVPMs. In the
rest of the section, we will consider the real-valued and interval-valued
probability measure defined in example~\ref{ex1} and \ref{ex3},
and see how to compute the expectation value of different random variables.
To simplify the discussion later, we will only consider a random variable~$X:\Omega\rightarrow\mathbb{R}$
which is a step function, i.e., $X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$,
where $E_{i}\in\events$ and $\mathbf{1}_{E}$ is the indicator function
defined by 
\begin{equation} 
\mathbf{1}_{A}(x)=\cases{ 
1 & if $x\in A$ ;\\ 
0 & if $x\notin A$. 
}
\end{equation}

Given a random variable~$X$ and a probability measure~$\pmeas$,
the expectation value of $X$ is 
\begin{eqnarray}
\int X\rmd\pmeas & = & \sum_{i=1}^{N}x_{i}\pmeas\left(E_{i}\right)\textrm{ .}\label{eq:expectation}
\end{eqnarray}
Intuitively, the expectation value is weighted average of the possible
outcomes of $X$ with the weight equal to the probability of each
outcome could happen. For example, consider the random variable~$X_{0}$
representing the number of tails. For the probability measure~$\pmeas$
in example~\ref{ex1}, the expectation value of $X_{0}$ is
\begin{eqnarray*}
\int X_{0}\rmd\pmeas & = & 0\cdot\pmeas(\{HH\})+1\cdot\pmeas(\{HT,TH\})+2\cdot\pmeas(\{TT\})\\
 & = & 0\cdot\frac{1}{3}+1\cdot\frac{2}{3}+2\cdot0=\frac{2}{3}\textrm{ .}
\end{eqnarray*}

There are various different integral defined for IVPMs, and the most
famous one is called the Choquet integral~\cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}.
Let $X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$ be a positive random
variable with $0\le x_{1}\le x_{2}\le\ldots\le x_{N-1}\le x_{N}$
and $\bar{\mu}:\events\rightarrow\mathscr{I}$ a classical IVPM such
that $l:\events\rightarrow\left[0,1\right]$ and $r:\events\rightarrow\left[0,1\right]$
be the left-end and the right-end of $\bar{\mu}$, respectively, i.e.,
$\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$,
the Choquet integral of $X$ with respect to $\bar{\mu}$ is 
\begin{eqnarray}
\int X\rmd\bar{\mu} & = & \sum_{i=1}^{N}\left(x_{i}-x_{i-1}\right)\bar{\mu}\left(\bigcup_{j\ge i}E_{j}\right)\textrm{ ,}\\
 & = & \sum_{i=1}^{N}x_{i}\left[l\left(\bigcup_{j\ge i}E_{j}\right)-l\left(\bigcup_{j\ge i+1}E_{j}\right),r\left(\bigcup_{j\ge i}E_{j}\right)-r\left(\bigcup_{j\ge i+1}E_{j}\right)\right]\label{eq:interval-expectation}
\end{eqnarray}
where $x_{0}=0$, and $x[l,r]=[xl,xr]$ for $x\ge0$. Similar to equation~(\ref{eq:expectation}),
equation~(\ref{eq:interval-expectation}) can be considered as weighted
average of the possible outcomes of $X$ as well, where the ``weight''
\[
\left[l\left(\bigcup_{j\ge i}E_{j}\right)-l\left(\bigcup_{j\ge i+1}E_{j}\right),r\left(\bigcup_{j\ge i}E_{j}\right)-r\left(\bigcup_{j\ge i+1}E_{j}\right)\right]\subseteq\left[l\left(E_{i}\right),r\left(E_{i}\right)\right]=\bar{\mu}\left(E_{i}\right)\textrm{ .}
\]
is tighter than the natural weight. This is reasonable because only
one event can happen at once so we don't be so pessimistic to accumulate
all possible fuzziness together, and use $\sum_{i=1}^{N}x_{i}\bar{\mu}\left(E_{i}\right)$.
